Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-01-30 draft.tex
add more aggrigated explanation results

	methods of synthetic audio detection. The surrogate model demonstrated strong performance on
	the dataset, even in comparison to previous methods. However, the explanations generated by
	this method do not seem to offer potential for useful interpretation using the perceptible
	features, based on an evaluation of aggregated local explanations using the proposed metrics.
	I believe, however, that the proposed method offers a promising basis for future research into
	explanations using perceptible features.

@@

					\[ \frac{ \frac{1}{N-1}\sum_{i=1}^{N-1}|T_{i}(\frac{1}{x}\sum_{n=i-m}^{i+m}T_{n})|}

@@

					\[ \frac{ \frac{1}{N-1}\sum_{i=1}^{N-1}|A_{i}(\frac{1}{x}\sum_{n=i-m}^{i+m}A_{n})|}

@@

			\begin{tabular}{c | c | c | c}
				Terminus & Training & Accuracy & EER \\

@@

				Simple & until 3474 & 95.02\% & 0.03702 \\
				3 hidden layers & until 10000 & 96.27\% & 0.03214 \\
				Simple & until 23833 2 epochs & 90.07\% & 0.90069 \\
				Simple & until 10000 2 epochs & 94.43\% & 0.04408 \\
				Simple & until 10000 & 94.28\% & 0.04840 \\
				3 convolutional layers & until 10000 & 62.80\% & 0.37198

@@

			\label{table:exp-results}
			\caption{Summary of the explanation results using both of the proposed features}

@@

		method, I have used the best performing model and generated explanations on the first 500
		samples of the testing data. I then summarized the explanations generated using both of
		the previously introduced metrics. A summary of the results of both metrics summarized by
		the feature is presented in table \ref{table:exp-results}. Measuring by both metrics, the
		MFCCs had the most positive and the most correct influence on the classification results of
		the model, With the jitter features as a distant second for trust and importance. This
		indicates that the MFCCs had not only the most influence, but also the most correct
		influence on the result of the classification, while the other features had either little,
		no or negative influence on the classification.

@@

		Because of this model's inclusion of several perceptible features, I made the hypothesis

@@

		table \ref{table:exp-results}, the perceptible features did not contribute as much to the
		end result of the classification as their imperceptible counterparts, in spite of their
		increased presence in the overall set of input features. Two features, the HNRs and the
		mel-spectrograms even had a net-negative impact on the correctness of the classifications
		in the experiments, leaving the question open, if future experiments may perform better
		without these features. Four perceptible features, f0 lengths, onset strengths,

@@

		between these features and the realness of audio samples in the dataset. In spite of this,
		I believe that it is still worth investigating if other kinds of perceptible features will
		have more of an influence on the overall classification or if there are other
		imperceptible features that could be used in the place of MFCCs that may provide a better
		balance between classification correctness and a proportional influence on the end result.

>>>>

	methods of synthetic audio detection. Several of the surrogate models demonstrated strong
	performance on the dataset, even in comparison to previous methods. However, the explanations
	generated by this method do not seem to offer much potential for useful interpretation using the
	perceptible features, based on an evaluation of aggregated local explanations using the
	proposed metrics, as the imperceptible features have a much larger influence on the
	classification result. I believe, however, that the proposed method offers a promising basis
	for future research into explanations using perceptible features.

@@

					\[ \frac{ \frac{1}{N-1}\sum_{i=1}^{N-1}|T_{i}
						(\frac{1}{x}\sum_{n=i-m}^{i+m}T_{n})|}

@@

					\[ \frac{ \frac{1}{N-1}\sum_{i=1}^{N-1}|A_{i}
						(\frac{1}{x}\sum_{n=i-m}^{i+m}A_{n})|}

@@

			\centering\scriptsize
			\begin{tabular}{c | c | c | c | c}
				Terminus & Features & Training & Accuracy & EER \\

@@

				Simple & standard & until 3474 & 95.02\% & 0.03702 \\
				3 hidden layers & standard & until 10000 & 96.27\% & 0.03214 \\
				Simple & standard & until 23833 2 epochs & 90.07\% & 0.90069 \\
				Simple & standard & until 10000 2 epochs & 94.43\% & 0.04408 \\
				Simple & standard & until 10000 & 94.28\% & 0.04840 \\
				3 convolutional layers & standard & until 10000 & 62.80\% & 0.37198 \\
				Simple & expand pitch-flucs & until 10000 & 93.31 & 0.05661

@@

			\label{table:exp-results-cterm}
			\caption{Summary of the aggregated explanation results of the best performing model}
				HNRs & 0.1072 & -0.0148 \\
				mel spectrogram & 0.0428 & -0.0187 \\
				MFCC & 0.4383 & 0.2345 \\
				f0 lengths & 0.0340 & -0.0057 \\
				onset strengths & 0.0101 & -0.0027 \\
				intensities & 0.0021 & 0.0021 \\
				pitch fluctuations & 0.0059 & 0.0020 \\
				jitter features & 0.2439 & 0.0019 \\
				shimmer features & 0.1664 & 0.0070
			\label{table:exp-results-more-pitch-flucs}
			\caption{Summary of the aggregated explanation results of the model with expanded
			pitch fluctuation features}

@@

		method, I have used selected two models from the last section and generated explanations
		on the first 500 samples of the testing data. I then summarized the explanations generated
		using both of the previously introduced metrics. A summary of the results of both metrics
		summarized by the feature is presented in tables \ref{table:exp-results-cterm} and
		\ref{table:exp-results-more-pitch-flucs}. 
		\subsubsection{Best-Performing Model}
		Measuring by both metrics, the MFCCs had the most positive and the most correct influence
		on the classification results of the model, With the jitter features as a distant second
		for trust and importance. This indicates that the MFCCs had not only the most influence,
		but also the most correct influence on the result of the classification, while the other
		features had either little, no or negative influence on the classification.
		\subsubsection{Model with Expanded Pitch-Fluctuation Features}
		This model had similarly unpromising results, with the main difference being that all
		metrics have non-zero values, meaning that all the present features had an influence on
		the classifications which were used to generate the explanations. As with the previous
		model, HNRs and mel-spectrograms had negative overall trust values, with the addition of
		the onset strengths also presenting a negative value, implying that these features have
		caused more harm than good in the tested classifications.

@@

		Because of these models' inclusion of several perceptible features, I made the hypothesis

@@

		the tables \ref{table:exp-results-cterm} \ref{table:exp-results-more-pitch-flucs}, the
		perceptible features did not contribute as much to the end result of the classification as
		their imperceptible counterparts, in spite of their increased presence in the overall set
		of input features. Two features, the HNRs and the mel-spectrograms even consistently had a
		net-negative impact on the correctness of the classifications in the experiments, leaving
		the question open, if future experiments may perform better without these features. In the
		case of the best performing model, four perceptible features, f0 lengths, onset strengths,

@@

		between these features and the realness of audio samples in the dataset. These features
		did have an impact on the performence of the model with expanded pitch fluctuations, but
		the difference was still minimal. In spite of this, I believe that it is still worth
		investigating if other kinds of perceptible features will have more of an influence on the
		overall classification or if there are other imperceptible features that could be used in
		the place of MFCCs that may provide a better balance between classification correctness
		and a proportional influence on the end result.
